On the power of oritatami cotranscriptional folding with unary bead sequence\thanks{This work is supported in part by KAKENHI Grant-in-Aid for Challenging Research (Exploratory) No.~18K19779 and JST Program to Disseminate Tenure Tracking System No.~6F36 granted to S.~Z.~F. and S.~S.}

### 2018-12-05 arity1.tex

\subsubsection{\ref{sec:unary}.2 \ \ $\delta = 3$\\}

\begin{enumerate}
\item Let us discuss that a fixed bead on a side of $\hexagon_p^n$ has a free hand, and bead $i$ binds to it. If the most stable conformation formed by beads $i, i+1, i+2$ has only two hydrogen bonds, it will be nondeterministic because there are at least two possibilities, see Fig.~\ref{fig:beadi}. Therefore, it needs to make three hydrogen bonds to deterministically stabilize $i$.\\

@@

There are the five cases in which beads $i, i+1, i+2$ can have three bonds, see Fig.~\ref{fig:3bonds}. There are beads having a free hand at (1,0) and (-1,0) to realize (b) and (e) in Fig.~\ref{fig:3bonds} before bead $i-1$ is fixed at (0,0). One of these two beads may be a predecessor of bead $i-1$, but at least one bead is not. A bead $i-1$ makes a hydrogen bond with one of these two beads which is not its predecessor when it is fixed at (0,0). That is, it is impossible to put three beads having a free hand at (-1,0), (0,0) and (1,0) when a bead $i$ is fixed, and then, to realize (b) and (e) in Fig.~\ref{fig:3bonds}. (d) in Fig.~\ref{fig:3bonds} becomes nondeterministic when bead $i$ is fixed because bead $i$ can be fixed at (-1,0) and bond with (-2,0), bead $i+1$ can be placed (0,1) and bond with (0,0) and bead $i+2$ can be placed (1,1) and bond with (0,1). If it makes three bonds once, such as (a) and (c) in Fig.~\ref{fig:3bonds}, it will need to make three bonds forever to be deterministic. Similarly to case 1. in the previous section, (a) and (c) in Fig.~\ref{fig:3bonds} become nondeterministic finally when it reaches a first corner of $\hexagon_p^n$.

\item Let us discuss that bead $i$'s predecessor at (0,0) has no free hand. Let us assume that there is a bead at (1,0) which has a free hand. If there is a bead at (1,0) and beads $i, i+1, i+2$ make only two hydrogen bonds, it will become nondeterministic as in Fig.~\ref{fig:beadat(1,0)}. That is, beads $i, i+1, i+2$ need to make three bonds to stabilize $i$. If they make three hydrogen bonds, the situation is analogous to (a) and (c) in Fig.~\ref{fig:3bonds}. There are not any bead having a free hand at (1,0) or (-1,0) to prevent from being (a) and (c) in Fig.~\ref{fig:3bonds}. Let us discuss the moment after bead $i$ is fixed outside $\hexagon_p^n$. Now bead $i$ has a free hand because there are no beads having a free hand at (1,0) or (-1,0). If beads $i+1, i+2, i+3$ can form only two bonds, it will be nondeterministic because there are at least two possible such conformations, as in Fig.~\ref{fig:beadi2}. Hence, they need to form three bonds to deterministically stabilize, such as in Fig.~\ref{fig:3bonds2}. (a) in Fig~\ref{fig:3bonds2} becomes (a) in Fig.~\ref{fig:3bonds}. (b) in Fig.~\ref{fig:3bonds2} becomes nondeterministic when bead $i$ is fixed because bead $i$ can be fixed at (1,0).\\

@@

It can yield only finite conformation and the length is $3n^2+n^2+1$.

>>>>

\subsubsection{\ref{sec:unary}.2 \ \ $\delta = 3$\\}
As before, let us assume that the last bead stabilized within $\hexagon_p^n$ is at point $(0,0)$ and the next bead is the first to be stabilized outside $\hexagon_p^n$ at point $(1,1)$. Depending on whether bead $i-1$ at point (0,0) has a free hand to form a bond or not, we distinguish two cases.
\begin{enumerate}
\item  Bead $i-1$ at (0,0) has a free hand.\\ If the most stable conformation formed by beads $i, i+1, i+2$ adds only two  bonds, it will be nondeterministic because there are at least two possibilities, see Fig.~\ref{fig:beadi}. Therefore, it needs to make three new bonds to deterministically stabilize $i$.\\

@@

There are five possible cases in which beads $i, i+1, i+2$ can add three bonds, see Fig.~\ref{fig:3bonds}. In the cases (b) and (e) in Fig.~\ref{fig:3bonds}, there are beads having a free hand at (1,0) and (-1,0) already stabilized before bead $i-1$ is fixed at (0,0). One of these two beads may be a predecessor of bead $i-1$, but at least one of them is not. When bead $i-1$ is fixed at (0,0), it makes a bond with the one of these two beads which is not its predecessor. This means that it is impossible to have three beads at (-1,0), (0,0) and (1,0), each with a free hand, when bead $i$ is fixed, and consequently, cases (b) and (e) in Fig.~\ref{fig:3bonds} cannot occur. Case (d) in Fig.~\ref{fig:3bonds} becomes nondeterministic when bead $i$ is fixed because bead $i$ can be fixed at (-1,0) and bond with (-2,0), bead $i+1$ can be placed (0,1) and bond with (0,0) and bead $i+2$ can be placed (1,1) and bond with (0,1). If it makes three bonds once, such as (a) and (c) in Fig.~\ref{fig:3bonds}, it will need to make three bonds forever to be deterministic. Similarly to case 1. in the previous section, cases (a) and (c) in Fig.~\ref{fig:3bonds} lead to nondeterministism eventually, when the transcript reaches a first corner of $\hexagon_p^n$.

\item Bead $i-1$ at (0,0) does not have a free hand.\\
(i) First, let us assume that there is a bead at (1,0) which has a free hand. If there is a bead at (1,0) and beads $i, i+1, i+2$ add only two bonds, we instantly get nondeterministism as in Fig.~\ref{fig:beadat(1,0)}. Hence, beads $i, i+1, i+2$ need to make three bonds to stabilize $i$, but if they make three hydrogen bonds, the situation is analogous to (a) and (c) in Fig.~\ref{fig:3bonds}.

%There are not any bead having a free hand at (1,0) or (-1,0) to prevent from being (a) and (c) in Fig.~\ref{fig:3bonds}. 
(ii) Now consider when there is no bead at (1,0) or there is one, but with no free hand. Let us discuss the moment after bead $i$ is fixed outside $\hexagon_p^n$. Now bead $i$ has a free hand because it cannot bind to (1,0). If beads $i+1, i+2, i+3$ can form only two bonds, it will be nondeterministic because there are at least two possible such conformations, as in Fig.~\ref{fig:beadi2}. Hence, they need to form three bonds to deterministically stabilize, such as in Fig.~\ref{fig:3bonds2}. Case (a) in Fig.~\ref{fig:3bonds2}. becomes the same as case (a) in Fig.~\ref{fig:3bonds}. Case (b) in Fig.~\ref{fig:3bonds2} becomes nondeterministic already when bead $i$ is fixed because bead $i$ could also be fixed at (1,0).\\

@@

We have shown that at $\alpha=1$ and  $\delta=3$, unary oritatami systems can only fold finite length transcripts deterministically, and the length of that transcript is bounded by the size of the regular hexagon of radius $n+1$ surrounding the first bead of the transcript. This gives us the upper bound $|\hexagon_p^{n+1}|=3(n+1)^2+(n+1)^2+1$,  where $n$ is the length of the seed.

### 2018-12-05 paper.tex

\title{On the power of oritatami cotranscriptional folding with unary bead sequence\thanks{This work is supported in part by KAKENHI Grant-in-Aid for Challenging Research (Exploratory) No.~18K19779 granted to S.~Z.~F. and S.~S. and JST Program to Disseminate Tenure Tracking System No.~6F36 granted to S.~S.}
}
\titlerunning{Power of unary oritatami cotranscriptional folding}
\author{
Szil\'{a}rd Zsolt Fazekas\inst{1} \and 
Kohei Maruyama \and
Reoto Morita \and \\
Shinnosuke Seki\inst{2}\thanks{Corresponding author (\email{s.seki@uec.ac.jp})}}
\institute{
Akita University, 
Graduate School of Engineering Science, 
1-1 Tegate Gakuen-machi, Akita, 0108502, Japan \\
%\email{szilard.fazekas@ie.akita-u.ac.jp}
\and
The University of Electro-Communications, 
Graduate School of Informatics and Engineering, 
1-5-1 Chofugaoka, Chofu, Tokyo, 1828585, Japan \\
}

@@

\begin{proposition}\label{prop:check_validity}
	For any rule set $R$, arity $\alpha$ and conformation $C = (P,w,H)$ it is possible to check whether $C$ is $R$-valid and whether $C\in \mathcal{C}_{\leq \alpha}$ in time $\mathcal{O}(|H|\cdot|w|\cdot|R|)$.
\end{proposition}
\begin{proof}
	To check whether $C$ is $R$-valid:
	\begin{enumerate}
		\item FOR each $(i,j)\in H$:
		\item \hspace{1cm} IF $(w[i],w[j])\notin R$ THEN answer NO and HALT
		\item answer YES and HALT
	\end{enumerate}	
	Checking the condition in 2. can be done in $\mathcal{O}(|w|\cdot|R|)$ time for any reasonable representation of $w$ and $R$, hence the whole process takes $\mathcal{O}(|H|\cdot |w|\cdot|R|)$ time.	
	To check the arity constraint $C\in \mathcal{C}_{\leq \alpha}$: 
	\begin{enumerate}
		\item FOR each $i\in \{1,\dots,|w|\}$:
		\item \hspace{1cm} IF $\mathrm{degree}(i)=|\{j | (i,j)\in H \}|>\alpha$ THEN answer NO and HALT
		\item answer YES and HALT
	\end{enumerate}	
	Checking the condition in 2. can be done in $\mathcal{O}(|H|)$ time for any reasonable representation of $H$, hence the whole process takes $\mathcal{O}(|w|\cdot|H|)$ time.
\end{proof}
\begin{theorem}\label{thm:OS_to_2dTM}
	There is an algorithm that simulates any oritatami system $\Xi = (\Sigma, R, \delta, \alpha, \sigma, w)$ in time $2^{\mathcal{O}(\delta)}\cdot |R|\cdot|w|$. 
\end{theorem}
\begin{proof}
	Take any step in the computation, up to which some $i \ge 0$ first beads of $w$ have been stabilized, with the last bead at a point $p$. 
	The number of all possible elongations of the current conformation by the next $\delta$-beads is $(6 \times 5^{\delta-1}) \times ((2^4)^{\delta-1} \times 2^5) \in 2^{O(\delta)}$. 
	By Proposition~\ref{prop:check_validity}, we can check for each of these elongations whether its arity is at most $\alpha$ or not and whether it is $R$-valid or not in time $\mathcal{O}((2^4)^{\delta-1}\cdot2^5\cdot \delta\cdot|R|)=2^{\mathcal{O}(\delta)}\cdot|R|$.  Therefore, the total running time is $2^{\mathcal{O}(\delta)}\cdot |R|\cdot|w|$.
\end{proof}

\begin{corollary}
	For fixed $\delta$ and $\alpha$, the class of problems solvable by oritatami systems $(\Sigma, R, \delta, \alpha, \sigma, w)$ is included in $\mathrm{DTIME}(n^3)$.
\end{corollary}
\begin{proof}
	The claim follows from Theorem~\ref{thm:OS_to_2dTM} and the fact that $|R|$ is implicitly bounded by $|w|^2$.
\end{proof}

>>>>

\title{On the power of oritatami cotranscriptional folding with unary bead sequence\thanks{This work is supported in part by KAKENHI Grant-in-Aid for Challenging Research (Exploratory) No.~18K19779 granted to S.~Z.~F. and S.~S. and JST Program to Disseminate Tenure Tracking System No.~6F36 granted to S.~S.}
}
\titlerunning{Power of unary oritatami cotranscriptional folding}
\author{
Szil\'{a}rd Zsolt Fazekas\inst{1} \and 
Kohei Maruyama\thanks{Primary corresponding author (\email{k.maruyama@uec.ac.jp})} \and
Reoto Morita \and \\
Shinnosuke Seki\inst{2}\thanks{Secondary corresponding author (\email{s.seki@uec.ac.jp})}
}
\institute{
Akita University, 
Graduate School of Engineering Science, 
1-1 Tegate Gakuen-machi, Akita, 0108502, Japan \\
%\email{szilard.fazekas@ie.akita-u.ac.jp}
\and
The University of Electro-Communications, 
Graduate School of Informatics and Engineering, 
1-5-1 Chofugaoka, Chofu, Tokyo, 1828585, Japan \\
}

@@

\begin{proposition}\label{prop:check_validity}
	For any rule set $R$, arity $\alpha$ and conformation $C = (P,w,H)$ it is possible to check whether $C$ is $R$-valid and whether $C\in \mathcal{C}_{\leq \alpha}$ in time $\mathcal{O}(|H|\cdot|w|\cdot|R|)$.
\end{proposition}
\begin{proof}
	To check whether $C$ is $R$-valid:
	\begin{enumerate}
		\item FOR each $(i,j)\in H$:
		\item \hspace{1cm} IF $(w[i],w[j])\notin R$ THEN answer NO and HALT
		\item answer YES and HALT
	\end{enumerate}	
	Checking the condition in 2. can be done in $\mathcal{O}(|w|\cdot|R|)$ time for any reasonable representation of $w$ and $R$, hence the whole process takes $\mathcal{O}(|H|\cdot |w|\cdot|R|)$ time.	
	To check the arity constraint $C\in \mathcal{C}_{\leq \alpha}$: 
	\begin{enumerate}
		\item FOR each $i\in \{1,\dots,|w|\}$:
		\item \hspace{1cm} IF $\mathrm{degree}(i)=|\{j | (i,j)\in H \}|>\alpha$ THEN answer NO and HALT
		\item answer YES and HALT
	\end{enumerate}	
	Checking the condition in 2. can be done in $\mathcal{O}(|H|)$ time for any reasonable representation of $H$, hence the whole process takes $\mathcal{O}(|w|\cdot|H|)$ time.
\end{proof}

\begin{theorem}\label{thm:OS_to_2dTM}
	There is an algorithm that simulates any deterministic oritatami system $\Xi = (\Sigma, R, \delta, \alpha, \sigma, w)$ in time $2^{\mathcal{O}(\delta)}\cdot |R|\cdot|w|$. 
\end{theorem}
\begin{proof}
	Take any step in the computation, up to which some $i \ge 0$ first beads of $w$ have been stabilized, with the last bead at a point $p$. 
	The number of all possible elongations of the current conformation by the next $\delta$-beads is $(6 \times 5^{\delta-1}) \times ((2^4)^{\delta-1} \times 2^5) \in 2^{O(\delta)}$. 
	By Proposition~\ref{prop:check_validity}, we can check for each of these elongations whether its arity is at most $\alpha$ or not and whether it is $R$-valid or not in time $\mathcal{O}((2^4)^{\delta-1}\cdot2^5\cdot \delta\cdot|R|)=2^{\mathcal{O}(\delta)}\cdot|R|$.  Therefore, the total running time is $2^{\mathcal{O}(\delta)}\cdot |R|\cdot|w|$.
\end{proof}

\begin{corollary}
	For fixed $\delta$ and $\alpha$, the class of problems solvable by deterministic oritatami systems $(\Sigma, R, \delta, \alpha, \sigma, w)$ is included in $\mathrm{DTIME}(n^3)$.
\end{corollary}
\begin{proof}
	The claim follows from Theorem~\ref{thm:OS_to_2dTM} and the fact that $|R|$ is implicitly bounded by $|w|^2$.
\end{proof}
